\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-08-15 popl-2019-journal-v/proving-full-abstraction.tex
Minor typesetting improvements.
Typesetting fix.

In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
\begin{lemma}[$\blater$ c.o.f.e.]
  \label{def:later-cofe}
  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
  \]
  where
  \[
    x \nequal[n]_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{}\\
        x \nequal[n-1]_X x'
      \end{array}
    \right.
  \]
\end{lemma}
The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.

@@

In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
As a reminder, we sketched the memory invariants to as follows
\[
  \World \monfun \Rel{\MemSeg \times \MemSeg}
\]
To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
We do not need a construction for the world as it will be given as the solution to the recursive equation.
Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
%% URel
\begin{definition}[Uniform relation]
  \label{def:uniform-rel-cofe}
  Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
  \[
    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
  \]
  The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
  \[
    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
  \]
  where erasure is defined as
  \[
    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
  \]
  to get a preordered c.o.f.e.\, we use subset as the preorder.
\end{definition}
The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
This means that everything is equal at step $0$ as everything is left out.

Next, we define a c.o.f.e. of monotone, non-expansive functions
\begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
\label{def:monnefun-cofe}
  Given a preordered c.o.f.e. $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where 
  \[
    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
  \]
\end{definition}
In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\, and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.

>>>>

In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
\begin{lemma}[$\blater$ c.o.f.e.]
  \label{def:later-cofe}
  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
  \]
  where
  \[
    x \nequal[n]_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \text{ or }\\
        x \nequal[n-1]_X x'
      \end{array}
    \right. \qedhere
  \]
\end{lemma}
The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.

@@

In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
As a reminder, we sketched the memory invariants to as follows
\[
  \World \monfun \Rel{\MemSeg \times \MemSeg}
\]
To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
We do not need a construction for the world as it will be given as the solution to the recursive equation.
Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
%% URel
\begin{definition}[Uniform relation]
  \label{def:uniform-rel-cofe}
  Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
  \[
    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
  \]
  The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
  \[
    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
  \]
  where erasure is defined as
  \[
    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
  \]
  to get a preordered c.o.f.e.\ we use subset as the preorder.
\end{definition}
The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
This means that everything is equal at step $0$ as everything is left out.

Next, we define a c.o.f.e.\ of monotone, non-expansive functions
\begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
\label{def:monnefun-cofe}
  Given a preordered c.o.f.e. $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where 
  \[
    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
  \]
\end{definition}
In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\  and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
